\chapter{Introduction}
\acrfull{les} is a technique used to accurately simulate complex fluid flows, of which the Earth's atmosphere is an example. Specifically, \acrshort{les} explicitly resolves turbulence, as opposed to traditional numerical weather and climate models, which suffer from significant uncertainties related to parameterizations of turbulent processes \citep{schalkwijkWeatherForecastingUsing2015}. However, due to resolution requirements, \acrshort{les} is computationally too expensive to apply on large domains. 

Since it was discovered that \acrfullpl{gpu} can be used to accelerate \acrfull{ai} models, the complexity of these models has been increasing at an ever-increasing rate \citep{mittalSurveyTechniquesOptimizing2019}. This has sparked the onset of a feedback loop, where successes in \acrshort{ai} lead to improvements in \acrshort{gpu} technology, and vice versa. \acrshort{ai} models consist of a large number of parallel computations, an area where the \acrshort{gpu} excels. Fortunately, \acrshort{les} models are also comprised of such computations. Therefore, \acrshortpl{gpu} can help accelerate \acrshort{les} models and enable simulations on larger domains.

This thesis describes the process of \acrshort{gpu}-accelerating \acrshort{dales}, an \acrshort{les} code tailored for atmospheric flows, by using the \acrshort{acc} programming model. The efforts are limited to the dynamical core and moist thermodynamic routines of \acrshort{dales}. Furthermore, the model is going to be adapted such that multiple \acrshortpl{gpu} can be used simultaneously. Validation of the accelerated model will be done through an approach based on ensemble statistics. A study of the performance of the model will be done on computer systems on two different scales: a desktop system with consumer hardware, and the Snellius supercomputer. For both systems, the speedup of the accelerated model is going to be examined and on Snellius, the scaling to multiple \acrshortpl{gpu} will be tested.

The outline of this report is as follows: in \Autoref{chap:les} the theory behind the modeling of turbulent flows is discussed and a technical description of \acrshort{dales} is given. \Autoref{chap:gpgpu} will explain \acrshort{gpu} technology, and how \acrshortpl{gpu} can be used to accelerate fluid simulation codes. Then in \Autoref{chap:implementation}, the implementation of \acrshort{acc} in \acrshort{dales} will be described. \Autoref{chap:val_bench} will present the results of the validation and performance studies. Finally, conclusions and recommendations for future work are to be found in \Autoref{chap:conclusion_rec}.